Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-13 doc/report/msc-report.tex
separated data cubes etc into chapter: Concepts.. moved section to bottom.
report: added graphichx package introduction introduces aggregation introduction introduces dimensions by way of Aerodyne Corporation example freshened the goal; we should move parts of it into analysis, other parts to scope. added CONCEPTS IN DATA WAREHOUSING (briefly)

\documentclass[a4paper]{report}

@@

Each log record typically contains a number of fields. For example, a log of
visits to a web site might contain a timestamp, information about what country
the visitor is from and what web browser she used. The values of the fields
can be thought of as points along an axis; Sweden, Norway and Denmark are all
points along the country axis, for instance. This is an important concept and
it will be touched upon repeatedly throughout the text.

When the records in the data set have been analyzed, one can draw conclusions
about what countries generate the most traffic to the site, which web browser
is most popular in a certain country, etc.

@@

It is suitable to store the data across many storage nodes in what is called a
distributed file system and to to exploit the parallel nature of the records
by breaking up the data sets in pieces and have each piece be processed
separately on a node in a large compute cluster.

@@

@@

The goal is to build a tool that enables its users to easily turn log files
into aggregates. The tool should provide a simple mechanism for data analysis
(import) and querying (export).

The data analysis will require the user to write a MapReduce map. The queries
will be performed via an export API for consumption by computer programs.

The queries will always be made over a time range and for a single unit. The
tool will support analytical functions such as drill-down, slicing, etc.

@@

@@

\section{Stop making speeches}
A data warehouse is a repository of an organization's electronically stored
data. Data warehouses are designed to facilitate reporting and analysis.
\cite{wikipedia}

The core actions of a data warehouse is to load, transform and extract data.
The classic thinking in modelling data warehouses is to divide data into facts
and dimensons. The facts are measures categorized by dimensions.

In OLAP\footnote{On-Line Analytical Processing}, which is a kind of data
warehouse, the data is thought of as being structured as a cube. The cube is
typically modelled in a realational database with fact tables and dimension
tables, upon which the system performs joins to extract data.
\cite{olap_solutions}

>>>>

\documentclass[a4paper]{report}
\usepackage{graphicx}

@@

Each log record typically contains a number of fields. For example, a log of
visits to a web site might contain a timestamp, information about what country
the visitor is from and what web browser she used.

When the records in the data set have been analyzed, one can draw conclusions
about what countries generate the most traffic to the site, which web browser
is most popular in a certain country, etc. Summaries of data are generally
called aggregates. To aggregate data, for example to compute the sum of
visitors for all European countries, is a common technique for speeding up
data lookup in data stores.

@@

It is suitable to to exploit the parallel nature of the records by breaking up
the data sets in pieces and to have each piece be processed separately on a
node in a large compute cluster. It is also suitable to store the data across
many storage nodes in what is called a distributed file system.

@@

% introducing variables and dimensions.

Organizations measure their operations by considering many variables. When
these variables are placed in a spreadsheet, which organizations are wont to
do, the variables are set on the x and y axes. Each axis represents a logical
grouping of variables. In the example of a log file of visits to a web site
above, country of origin is one such variable. A variable is generally called
a dimension in this paper.

The International Aerodyne Corporation, a leading exporter of woodworking
nails, uses a spreadsheet to keep track of the number of nails exported to
each country every month (see figure).

% image of a spreadsheet.
%\includegraphics[width=70mm]{pictars/aerodyne-nails.png}
%\imagecaptionplease{crates of nails exported}

In the figure above, the time dimension is set on the x-axis and the country
dimension is set on the y-axis. If the Aerodyne Corporation were to start
keeping track of the kind of nails they sell (pins, tacks, spikes, etc.) in
addition to the country to which the nails were sold, they would effectively
have added a new dimension to their data.

Many organizations work with data that have more than three dimensions. Such
data is hard to manage in a spreadsheet, and the organizations need
specialized tools to analyze and visualize their data.

@@

The goal is to build a data store that enables its users to perform
multi-dimensional analysis.

In addition, the secondary goal is to build a tool that allows users to easily
store their data sets in the data store. The tool should provide a simple – as
in minimalistic – mechanism for data analysis (import) and querying (export).

% is mapreduce part of the goal? hardly.
The data analysis will require the user to write a MapReduce map. The queries
will be performed via an export API for consumption by computer programs.

% these are more likely part of the scope.
The queries will always be made over a time range and for a single unit. The
tool will support analytical functions such as drill-down, slicing, etc.

@@

% CONCEPTS IN DATA WAREHOUSING, by clark de jean
\chapter{Concepts in Data Warehousing}

This chapter will serve as an introduction to various concepts in data
warehousing.

\section{Data Warehouses}

A data warehouse is a repository of an organization's electronically stored
data. Data warehouses are designed to facilitate reporting and analysis; the
core actions of a data warehouse are to load, transform and extract data.

\section{Data Warehouse implementations}

\subsection{OLAP}

OLAP, which is an acronym for Online analytical processing, is an approach to
quickly answer multi-dimensional analytical queries.

At the core of any OLAP system is the OLAP cube (also known as a
multidimensional cube or a hypercube). It consists of numeric facts called
measures which are categorized by dimensions.

The cube is typically modelled in a relational database with fact tables
and dimension tables, upon which the system performs joins to extract data.
\cite{olap_solutions}

@@
